---
title: "IPL Case Study"
author: "Group-IV"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: no
    toc: yes
    toc_float: yes
  fontsize: 14 pt
  mainfont: Times New Roman
---
<style type="text/css">

}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 34px;
  color: Black;
  font-family: "Times New Roman", Times, serif;
}
h2 { /* Header 2 */
    font-size: 30px;
  color: Black;
  font-family: "Times New Roman", Times, serif;
}
h3 { /* Header 3 */
  font-size: 24px;
  font-family: "Times New Roman", Times, serif;
  color: Black;
}
body {
    color: black;
    font-size: 18px;
    font-family: 'Times New Roman';
    }
code.r{ /* Code block */
    font-size: 12px;
}
</style>

```{css, echo=FALSE}
.bg-danger {
  background-color: #FFE2DF;
  border: 1px;
  font-weight: bold;
}
```
```{css, echo=FALSE}
.bg-success {
  background-color: #F5F5F5;
  border: 1px;
  font-weight: bold;
}
```
# **1. Importing Libraries**
```{r, warning = FALSE, class.source="bg-danger", class.output="bg-success", message = FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(forcats)      
library(plotly)       
```
**forcats** is for fct_reorder() funtion and **plotly** is for ggplotly() function. 

# **2. Importing datasets**
```{r, warning = FALSE, class.source="bg-danger", class.output="bg-success"}
ipl.df <- read.csv("https://raw.githubusercontent.com/Roshann-Rai/IPL_case_analysis/main/ipl_player_performance.csv")
head(ipl.df)
str(ipl.df)
```
The ipl.df has 11 variables and 70 rows. Except **Player** and **Team** variables other are numeric.

# **3. Data Cleaning**
## **3.1 Missing Data?**
```{r, warning = FALSE, class.source="bg-danger", class.output="bg-success"}
#checking if there are missing data
colSums(is.na(ipl.df))
```
It seems there are no missing data.

## **3.2 Removing some characters**
Looks like the the **Player** variable has observations with **<a0>** character which must be removed. **gsub()** function helps to replace **<a0>** by space.
```{r, warning = FALSE, class.source="bg-danger", class.output="bg-success"}
#Removing <a0> from the player names
ipl.df$Player <- gsub("<a0>", " ", ipl.df$Player)
head(ipl.df)
```

# **4. Data manipulation and data insights**
## **4.1 Creating new categorical variables**
```{r, warning=FALSE, class.source="bg-danger", class.output="bg-success"}
#Creating a new categorical variable from Avg variable
ipl.df1 <- ipl.df %>%
  mutate(Avg.run.category = case_when(Avg < 20 ~ "Less than 20",
                                      Avg < 40 & Avg >= 20 ~ "20-40",
                                      Avg < 60 & Avg >= 40 ~ "40-60",
                                      Avg < 80 & Avg >= 60 ~ "60-80",
                                      TRUE ~ "Above 80"),
         SR.category = case_when(SR < 75 ~ "Less than 75",
                                 SR < 150 & SR >= 75 ~ "75-150",
                                 SR < 225 & SR >= 150 ~ "150-225",
                                 TRUE ~ "Above 225"),
         player.category = case_when(Avg >= 28.58 & SR >= 133.12 ~ "Good Player Performance",
                                     TRUE ~ "Bad Player Performance"))
```

## **4.2 Summary Statistics**
Let's examine the summary statistics of Avg, SR and Salary variables.
```{r, warning=FALSE, class.source="bg-danger", class.output="bg-success"}
summary(ipl.df$Avg)
summary(ipl.df$SR)
summary(ipl.df$Salary)
```

## **4.3 Boxplots**
### **4.3.1 Boxplot of Strike rate of players**
```{r, warning=FALSE, fig.align="center", fig.dim=c(4,4), class.source="bg-danger", class.output="bg-success"}
boxplot(ipl.df$SR)
title("Strike Rate of Players")
```
It looks like there are 4 outliers in SR variable.

### **4.3.2 Boxplot of Salary of Players**
```{r, warning=FALSE, fig.align="center", fig.dim=c(4,4), class.source="bg-danger", class.output="bg-success"}
boxplot(ipl.df$Salary)
title("Salary of Players")
```

### **4.3.3 Boxplot of Average runs of players**
```{r, warning=FALSE, fig.align="center", fig.dim=c(4,4), class.source="bg-danger", class.output="bg-success"}
boxplot(ipl.df$Avg)
title("Average runs of players")
```
It looks like there are two outliers in this variable.

## **4.4 Outstanding Players**
```{r, warning=FALSE, class.source="bg-danger", class.output="bg-success"}
#Players with outstanding performance
players.no <- ipl.df %>%
  filter(Avg > 35.16 & SR > 148.47)

players.no
```

##  **4.5 Plots**
### **4.5.1 Barplot of average runs of all players**
```{r, warning=FALSE, fig.dim=c(8,10), fig.align = "center", class.source="bg-danger", class.output="bg-success", message=FALSE}
#plotting the players avg run
p1 <- ipl.df %>%
  mutate(player.reordered = fct_reorder(Player, Avg)) %>%
  ggplot(aes(player.reordered, Avg)) +
  geom_col() +
  #geom_text(aes(label = Avg), hjust = 0, size = 3.2) +
  labs(x = "Average runs by Player",
       y = "",
       title = "Players Average run") +
  theme_bw() +
  coord_flip()
ggplotly(p1)
```

### **4.5.2 Barplot of salary of all players**
```{r, warning=FALSE, fig.dim=c(8,10), fig.align = "center", class.source="bg-danger", class.output="bg-success", message=FALSE}
#plotting the players salary
p2 <- ipl.df %>%
  mutate(player.reordered = fct_reorder(Player, Salary)) %>%
  ggplot(aes(player.reordered, Salary)) +
  geom_col() +
  labs(x = "Salary in USD million",
       y = "",
       title = "Salary earned by players in IPL") +
  #geom_text(aes(label = Salary), hjust = 0, size = 3.2) +
  theme_bw() +
  coord_flip()
ggplotly(p2)
```

### **4.5.3 Scatter plot of salary vs average runs**
```{r, warning = FALSE, fig.align = "center", class.source="bg-danger", class.output="bg-success", message=FALSE}
#scatter plot of salary vs avg runs
p3 <- ipl.df1 %>%
  ggplot(aes(Avg, Salary)) +
  geom_point(aes(color = Avg.run.category)) +
  geom_smooth(method = "lm", se = F) +
  labs(x = "Average run by players",
       y = "Salary in USD million",
       title = "Scatter plot of Salary vs Average runs") +
  theme_bw() 
ggplotly(p3)
```

### **4.5.4 Scatter plot of salary vs strike rate**
```{r, warning = FALSE, fig.align = "center", message=FALSE, class.source="bg-danger", class.output="bg-success"}
#Scatter plot of salary vs Strike rate
p4 <- ipl.df1 %>%
  ggplot(aes(SR, Salary)) +
  geom_point(aes(color = SR.category)) +
  geom_smooth(method = "lm", se = F) +
  labs(x = "Strike Rate by players",
       y = "Salary in USD million",
       title = "Scatter plot of Salary vs strike rate") +
  theme_bw()
ggplotly(p4)
```

## **4.6 Karl Pearson Correlation coefficients**
```{r, warning = FALSE, class.source="bg-danger", class.output="bg-success", message = FALSE}
#Karl pearson correlation coefficients
cor.df <- ipl.df %>%
  select(Salary, SR, Avg) 
cor(cor.df)
```
The correlation coefficient between Salary and Strike rate is 0.1056 indicating very weak but positive relationship between them. Similarly, the correlation coefficient between Average runs and Strike rate is 0.4587 indicating positive and moderate strength of association between the two while the correlation coefficient between Salary and Average runs is 0.292 indicating weak and positive relationship between the two.

## **4.7 Two Sample t-test**
```{r, warning = FALSE, class.source="bg-danger", class.output="bg-success", message=FALSE}
#Two sample t test
t.test.df <- ipl.df1 %>%
  select(player.category, Salary) %>%
  group_by(player.category) %>%
  summarise(Mean = mean(Salary),
            SD = sd(Salary),
            samp_size = n())
t.test.df
```
The mean salary of player with bad performance is USD 0.77 million while that of player with good performance is USD 1.24 million. Since the mean salary of player with good performance is greater than player with bad performance, we will be evaluating if the player with bad performance is actually lesser than that of player with good performance using sample t-test.

 **H0:Mean salary of player with bad performance is greater than or equal to that of player with good performance.**<br>
 
 **H1: Mean salary of player with bad performance is less than that of player with good performance.**<br>
 
```{r, warning = FALSE, class.source="bg-danger", class.output="bg-success", message=FALSE}
t.test(Salary ~ player.category, ipl.df1, alternative = "less")
```
At 5% significance level, p-value is less than 0.05 so, alternative hypothesis is accepted.
i.e. Salary of players with good performance is greater than that of player with bad performance.

## **4.8 Linear Regression**
```{r, warning = FALSE, class.source="bg-danger", class.output="bg-success", message=FALSE}
#linear regression model of player average run
#Avg runs has been assumed as the performance indicator.
model.salary <- lm(Salary ~ Avg + SR, data = ipl.df)
summary(model.salary)
```
Salary is significantly affected Average run (at 5% level of significance) rate but not by strike rate.
The linear model is salary = 0.6421 + 0.014Avg + Error.
Adjusted R-square = 0.059 i.e., salary is only determined 5% by Avg variable and rest by the variables not present in this linear regression model.
It suggests that we need to identify more variables to determine the salary of player.

